---
title: "Economist"
author: "Hanh Nguyen"
date: "5/21/2017"
output:
  pdf_document: default
  word_document: default
---

The dataset (Economist.csv) consists of countries scored on how corrupt their public sectors are seen to be (Corruption Perceptions Index - CPI) and on achievement in key dimensions of human development: a long and healthy life, being knowledgeable and have a decent standard of living (Human Development Index - HDI).  


```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(reshape2)
dat = read.csv(file="/Users/user/GitHub/data-vis/dataSets/EconomistData.csv",header=TRUE)
dat$CPI.Rank = rank(dat$CPI,ties.method="first")
```

# 1. Visualization on national level  

## Human Development summary

```{r}
dat %>%
  ggplot(aes(x = Region, y = HDI)) + geom_boxplot()
```

## Human Development on the world map  

```{r}
dfworldmap = map_data("world")
map_data = merge(dfworldmap, dat,by.x="region",by.y="Country",all=TRUE)
map_data = transform(map_data, HDI = cut_interval(HDI, 10))
map_data = map_data[order(map_data$order),]
map_data %>%
  ggplot(aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = HDI)) +
  scale_fill_grey(start=1, end =.3) + 
  geom_path(colour='black') +
  coord_map() +
  ggtitle("Human Development Index Ranking of the world")
```

*cut_interval* makes n groups with equal range; *cut_number* makes n groups with (approximately) equal numbers of observations; *cut_width* makes groups of width.  

Some countries are named differently from dat and dfworldmap so renaming them is necessary.  

Countries in dat that do not have any match in dfworldmap: **df1 = map_data[is.na(map_data$long),]**  

Countries in dfworldmap that do not have any match in dfworldmap: **df2 = map_data[is.na(map_data$CPI),]** and **sort(unique(df2$region))**  

```{r}
dat$Country[dat$Country=="Korea (South)"] = "South Korea"
dat$Country[dat$Country=="United States"] = "USA"
dat$Country[dat$Country=="Britain"] = "UK"
dat$Country[dat$Country=="Trinidad and Tobago"] = "Trinidad"
dat$Country[dat$Country=="Congo  Republic"] = "Democratic Republic of the Congo"
```


## Corruption Perceptions  

```{r}
dat %>%
  ggplot(aes(x = Region, y = CPI)) + geom_boxplot()
```

## Corruption Perceptions on the world map  

```{r}
map_data = merge(dfworldmap, dat,by.x="region",by.y="Country",all.x=TRUE)
map_data = transform(map_data, CPI = cut_number(CPI, 5))
map_data = map_data[order(map_data$order),]
map_data %>%
  ggplot(aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = CPI)) +
  scale_fill_grey(start=1, end =.3) + 
  geom_path(colour='black') +
  coord_map() +
  ggtitle("Corruption Perceptions Index Ranking of the world")
```

## Corruption Perceptions and Human Development  

geom_smooth() is used to add a smooth line.

```{r}
dat %>%
  ggplot(aes(x = CPI, y = HDI)) + 
  geom_point(aes(col=Region),shape=1,stroke=1,size=2,alpha=.8) +
  geom_smooth(method="lm",se=FALSE,col="black") +
  labs(x="Corruption Perceptions Index",y="Human Development Index") +
  theme_classic() +
  geom_text_repel(data=subset(dat, HDI.Rank <=10 | CPI.Rank <=10),aes(CPI, HDI, label = Country),size=2.5) +
  expand_limits(x = 0, y = 0)
```
  
# 2. Visualization on regional level 

We first aggregate (group by) the data by region  

```{r}
reg_dat= dat %>%
  group_by(Region) %>%
  summarize(avgCPI=mean(CPI,na.rm=T),avgHDI=mean(HDI,na.rm=T)) %>%
  arrange(avgCPI, avgHDI)
reg_dat$CPI.Rank = rank(reg_dat$avgCPI,ties.method="first")
reg_dat$HDI.Rank = rank(reg_dat$avgHDI,ties.method="first")
```

## Corruption Perceptions and Human Development  

```{r}
reg_dat %>%
  ggplot(aes(x = avgCPI, y = avgHDI)) + 
  geom_point(aes(col=Region),size=5) +
  labs(x="Corruption Perceptions Index",y="Human Development Index") +
  theme_classic() +
  geom_text_repel(aes(avgCPI, avgHDI, label = Region),size=3) + 
  expand_limits(x = 0, y = 0) +
  guides(col=F)
```
Example: a lm for entire dataset and other lms for subsets

# Plot 3: include a lm for the entire dataset in its whole

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, linetype = 2)
```

 
  
Source:  
http://tutorials.iq.harvard.edu/R/Rgraphics/Rgraphics.html#org93999d8  
https://www.transparency.org/research/cpi/overview  
http://hdr.undp.org/en/content/human-development-index-hdi  


