---
title: "UNvotes"
author: "Hanh Nguyen"
output: pdf_document
---

```{r setup}
library(dplyr)
library(unvotes)
library(ggplot2)
library(tidyr)
library(purrr)
library(broom)
```

Erik Voeten "Data and Analyses of Voting in the UN General Assembly" Routledge Handbook of International Organization, edited by Bob Reinalda (published May 27, 2013).  
Below is three datasets in the package and their following columns:  
**1. un_votes** provides information on the voting history of the United Nations General Assembly. Contains one row for each country-vote pair.  
• rcid: The roll call id; it is the primary key used to join with tables un_roll_calls and un_roll_call_issues
• vote: Vote result as a factor of yes/abstain/no (The original data included cases where a country was absent or was not yet a member. In this dataset these were filtered out to include only votes of Yes, Abstain, and No)  
• country: Country name, by official English short name (ISO)
```{r}
head(un_votes)
unique(un_votes$vote)
```
**2. un_roll_calls** provies information on each roll call vote of the United Nations General Assembly.  
• rcid: The roll call id  
• session: Session number. The United Nations holds one session per year; these started in 1946  
• importantvote: Whether the vote was classified as important by the U.S. State Department report "Voting Practices in the United Nations". These classifications began with session 39  
• date: Date of the vote, as a Date vector  
• unres: Resolution code  
• amend: Whether the vote was on an amendment; coded only until 1985  
• para: Whether the vote was only on a paragraph and not a resolution; coded only until 1985  
• short: Short description  
• descr: Longer description  

```{r}
head(un_roll_calls)
```

**3. un_roll_call_issues** provides issue (topic) classifications of roll call votes of the United Nations General Assembly, with one row for each pair of a roll call vote and an issue describing that vote. Many votes had no topic, and some have more than one.  
• rcid: The roll call id; used to join with un_votes and un_roll_calls  
• short_name: Two-letter issue codes  
• issue: Descriptive issue name  

```{r}
head(un_roll_call_issues)
```

Further details about the package and datasets can be found here https://github.com/dgrtwo/unvotes or by ??unvotes

We want to know the voting pattern by each year and each country. Therefore, we will merge **un_votes** and **un_roll_calls** by *rcid*. Furthermore, we will create another field named *year* derived from the column *date*  

```{r}
df = merge(x=un_votes, y=un_roll_calls, by="rcid", all.x=TRUE)
df$year <- as.numeric(format(df$date,"%Y"))
head(df)
```

To see the voting pattern by year, we group the df by year using group_by() function
```{r}
by_year = df %>%
  group_by(year) %>%
  summarize(total=n(), percent_yes = mean(vote=="yes"))
head(by_year)
```

The data frame **by_year** is actually a time series and by looking at the visualization, we can see a trend over time
```{r}
ggplot(by_year, aes(year, percent_yes)) +
  geom_line()
```

Adding the geom_smooth() function
```{r}
ggplot(by_year, aes(year, percent_yes)) +
  geom_line() +
  geom_smooth()
```

A different perspective is to see voting patterns among countries.

```{r}
by_country = df %>%
  group_by(country) %>%
  summarize(total=n(), percent_yes = mean(vote=="yes"))
head(by_country)
```

We sort the data frame by the number of votes and the % of "yes" votes in the ascending order
```{r}
arrange(by_country, total)
arrange(by_country, percent_yes)
```

We can recognize that the country that voted least frequently, Zanzibar, had only 2 votes in the entire dataset, thus its percent_yes is not meaningful. For this reason, we will exclude countries with fewer than 100 votes in total.

```{r}
by_country %>%
  arrange(percent_yes) %>%
  filter(total >= 100)
```

Lastly, we want to summarize by both year and country, constructing a dataset that shows what fraction of the time each country votes "yes" in each year.

```{r}
by_year_country = df %>%
  group_by(year, country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == "yes"))
head(by_year_country)
```

Looking at the US data
```{r}
US_by_year = by_year_country %>%
  filter(country=="United States")
head(US_by_year)
```

```{r}
ggplot(US_by_year, aes(x=year,y=percent_yes)) +
  geom_line()
```

Plotting just one country at a time is interesting, but it'd be more insightful to compare trends between countries. 
Here we're interested in 8 most powerful countries in 2017 and see how their historical voting behaviors are. According to https://www.usnews.com/news/best-countries/power-full-list, the Power subranking is based on an equally weighted average of scores from five country attributes that related to a country's power: a leader, economically influential, politically influential, strong international alliances and strong military alliances. The Power subranking score had a 7 percent weight in the overall Best Countries ranking.  

Those countries are the United States, Russia, China, the UK, Germany, France, Japan and Israel.  

Before we get their data, we need to find the exact names used. I will demonstrate one example, suppose we look for Russia.

```{r}
subset(un_votes,country=="Russia")
```
It seems that Russia is not the exact name used in the dataset! 
```{r}
russia_find = grepl("Russia", un_votes$country)
russia_df = un_votes[russia_find,]
head(russia_df)
```

So the right name is Russian Federation! Running the same commands, we will be able to figure out the names for these 8 countries

```{r}
countries_8 <- c("United States", "Russian Federation", "China","United Kingdom", "Germany","France", "Japan","Israel")
countries_8_by_year = by_year_country %>%
  filter(country %in% countries_8)
```

```{r}
ggplot(countries_8_by_year, aes(x=year,y=percent_yes,color=country)) +geom_line()
```

However, this type of graph could be tough to read. The alternative way is faceting.

```{r}
ggplot(countries_8_by_year, aes(x=year,y=percent_yes,color=country)) + geom_line() + facet_wrap(~ country)
```

The legend seems redundant so we remove it
```{r}
ggplot(countries_8_by_year, aes(x=year,y=percent_yes,color=country)) + geom_line() + facet_wrap(~ country) + theme(legend.position="none")
```

China consistantly had very high percents of "yes" votes except a drop in 2014. Japan had a bit lower percents fluctuating around .5 and .75. The United States started at a similar level but steadily was voting more "no" or "abstain" and its "yes" percents of votes even lowered to .12 during 1980s. Israel's voting pattern also had a similar movement to the United States's.

Optional: Feel free to explore countries that you're interested in!

**MODELING**

# Linear regression on the United States
```{r}
US_fit = lm(percent_yes ~ year, data= US_by_year)
summary(US_fit)
```

# Fit models for 4 countries
```{r}
UK_by_year = by_year_country %>%
  filter(country=="United Kingdom")
UK_fit = lm(percent_yes ~ year, data= UK_by_year)

France_by_year = by_year_country %>%
  filter(country=="France")
France_fit = lm(percent_yes ~ year, data= France_by_year)

China_by_year = by_year_country %>%
  filter(country=="China")
China_fit = lm(percent_yes ~ year, data= China_by_year)
```

Tidy models and combine them together

```{r}
US_tidied = tidy(US_fit)
UK_tidied = tidy(UK_fit)
France_tidied = tidy(France_fit)
China_tidied = tidy(China_fit)
```


# Analysis of Resolution Type 
A different angle is to look at the types of resolutions. There are 6 issue types as below  

```{r}
unique(un_roll_call_issues$issue)
```

We want to know if countries have any preference or particular voting patern for any issue. First, we join two datasets un_votes and un_roll_call_issues using rcid 

```{r}
head(un_votes)
head(un_roll_call_issues)
df2 = merge(x=un_votes, y=un_roll_call_issues, by="rcid", all.x=TRUE)
head(df2)
```

Plotting the data frame to see the number of resolutions by issue

```{r}
df2 %>%
  ggplot(aes(x=issue)) +
  geom_bar(aes(fill=issue))
```

Adding features to enhance the look
```{r}
df2 %>%
  ggplot(aes(x=issue)) +
  geom_bar(aes(fill=issue)) +
  coord_flip() + 
  theme(legend.position="none") +
  ggtitle("The number of Resolution types") +
  ylab("")
```

From the chart, Palestinian conflict is the major concern of United Nation during 1946-2014, following by Human rights, Arms control and disarmament and Colonialism. Interestingly, economic development gets the least attention.

Another way to get a similar result is to group by the dataset by issue as follows:

```{r}
by_issue = df2 %>%
  group_by(issue) %>%
  summarize(total = n(), percent_yes = mean(vote == "yes"))

head(by_issue)
```

All resolutions but Human Rights has 80% and above concensus ("yes" votes). Human Rights Resolutions have 75% votes with "yes".

Now we want to look at the data not only by issue but also by country  
```{r}
by_issue_country = df2 %>%
  group_by(issue, country) %>%
  summarize(total = n(), percent_yes = mean(vote == "yes"))

head(by_issue_country)
```

Let's take US as an example to see the country's voting pattern on different issues

```{r}
US_by_issue = by_issue_country %>%
  filter(country=="United States")

US_by_issue
```

Making the plot

```{r}
US_by_issue %>%
  ggplot(aes(x=issue,y=percent_yes)) +
  geom_bar(stat="identity",aes(fill=issue)) +
  coord_flip() + 
  theme(legend.position="none") +
  ggtitle("The number of Resolution types") +
  ylab("") +
  facet_wrap(~ country)
```

The United States seems to disagree with most resolutions, lowest "yes" voting is to Palestinian conflict.  

Again, we want to see 8 countries at one view for comparison. 
```{r}
countries_8 <- c("United States", "Russian Federation", "China","United Kingdom", "Germany","France", "Japan","Israel")

countries_8_by_issue = by_issue_country %>%
  filter(country %in% countries_8)

countries_8_by_issue %>%
  ggplot(aes(x=issue,y=percent_yes)) +
  geom_bar(stat="identity",aes(fill=issue)) +
  coord_flip() + 
  theme(legend.position="none") +
  ggtitle("The number of Resolution types") +
  ylab("") +
  facet_wrap(~ country)
```

We want to use k-means clustering to divide countries into 2 groups: "yes" group and "no" group. First, we remove countries that have small number of votes because the percentage might not be meaningful with small sample sizes.

```{r}
# remove countries with total votes less than 10 for each period each issue
by_issue_country = by_issue_country %>%
  filter(total >= 100)
dff1 = subset(by_issue_country,select=-total)
dff1$issue[dff1$issue=="Arms control and disarmament"] = "Arms"
dff1$issue[dff1$issue=="Human rights"] = "Human"
dff1$issue[dff1$issue=="Nuclear weapons and nuclear material"] = "Nuclear"
dff1$issue[dff1$issue=="Palestinian conflict"] = "Palestinian"
dff1$issue[dff1$issue=="Economic development"] = "Economic"
dff1$issue[is.na(dff1$issue)] = "Other"

dff2 = dff1 %>%
  spread(issue,percent_yes)
head(dff2)
summary(dff2)
dff2 =na.omit(dff2)

set.seed(20)
kmc <- kmeans(dff2[,-1], centers=2, iter.max=1000)
kmc
```

There are 48 countries in group 1 and 138 countries in group 2. It shows that group 2 is pro- UN resolutions while group 1 appears to disagree more with the resolutions. There are no distinct difference among types of resolutions, group 2 consistently voted "yes" much more than group 1 across all kinds of resolutions.

```{r}
# append cluster assignment
dff2_cluster <- data.frame(dff2, kmc$cluster)
head(dff2_cluster)
dim(dff2_cluster)
countries_8_cluster = dff2_cluster %>%
  filter(country %in% countries_8)
countries_8_cluster
```

The result shows that China is classified in group 2 while the rest in group 1. This again confirms the charts above.
Out of 8 countries, only China seems to agree with United Nations resolutions as it voted "yes" around 90% of the times. While the remaing 7 countries seem not to favor United Nations resolutions.

Sources:  
https://github.com/dgrtwo/unvotes  
https://www.kaggle.com/karimkardous/vote-dynamics/code